\malg: a novel interest management algorithm for distributed simulations of MMOGs

### 2008-05-14 textos/dsrt/engaoi.tex
FULL

In the past few years, many works have been done trying to distribute the network support for massively multiplayer online games. One approach is the peer-to-peer model \cite{endo2005dam, hampel2006ppa, knutsson2004pps}, which has some issues concerning game state consistency among the participating peers, vulnerability to cheating, and increased bandwidth usage by each peer. There are some proposals that try to minimize these problems. One of them is presented in \cite{schiele2007rpp}, where it is suggested the division of the virtual simulated game environment into regions, and for each region, a peer is chosen to be its coordinator. The function of this peer will be manage the players' interest. It will check who really needs each state update, so each player does not receive every update packet. This way, the peers' bandwidth usage is reduced. However, it will be still higher than with a server performing all the broadcasts. In the peer-to-peer model, each player usually must send updates to more than one other player. Besides, the chosen peer must be trustable to be the region coordinator.

@@

In \cite{zou2001egt}, multicast-based interest management schemes are also considered. A comparison is made between cell-based grouping, where each group is associated with a cell of a grid that compose the virtual environment, and grouping based on objects, where there is a multicast group for each object. It was found that there is a tradeoff between the cost of multicast group formation control messages and the cost of the state updates themselves.

@@

In \cite{morse1996iml}, it is made a survey of systems that use the interest management technique, describing the purpose, scope and what criteria are used by the management scheme employed by each one of them. It is then presented a taxonomy of such systems, classifying them according to: communication model, filtering focus and domain of responsibility. The model may be \textit{unicast}, \textit{multicast} or \textit{broadcast}. The filtering focus refers to what characteristics are observed for each object to perform the filtering: they may be \textit{intrinsic}, such as the values of the object's attributes (e.g. the exact coordinates of its location), or \textit{intrinsic}, such as which multicast group it is subscribed to. Finally, the domain of responsibility assigned to a interest manager, which checks for whom each state is relevant, can be \textit{static} or \textit{dynamic}. For instance, if an interest manager is assigned to a fixed area of the virtual environment, its domain of responsibility is static, but if he controls an area that may increase or decrease in size, its domain of responsibility is dynamic.

@@

Taking into consideration that the multicast communication model is not widely supported in the Internet \cite{elsayed2004alm}, %for technical and marketing reasons as well,
 in this work we opted to follow the unicast model. Each broadcast consists then of a set of unicast transmissions, one for each destination. Moreover, intrinsic filtering and dynamic domain of resposibility were used, in order to increase the precision of the interest management and consequently achieve a greater reduction in the state update traffic \cite{morse1996iml}.

@@

This work is based on a virtual environment partitioned into regions, each managed by a different server. The regions are contiguous, exploring the locality of the players' avatars. Thus, avatars next to each other will probably be located in the same region and, therefore, their clients will tend to be connected to the same server, so that their interaction is faster (Figure \ref{partitions}). Situations in which two players interacting with each other are connected to different servers imply a higher bandwidth usage, as it is required some sort of negotiation between the servers to which the different players are connected, in order to maintain the states of the simulation identical in both servers. Also, the hop count of each message between the clients of these players is increased, since there is one more intermediary.

@@

% use section* for acknowledgement
%\section*{Acknowledgment}

%The authors would like to thank...

>>>>

In the past few years, many works have been done trying to distribute the network support for massively multiplayer online games. One approach is the peer-to-peer model \cite{crippa2007pps, hampel2006ppa, knutsson2004pps}, which has some issues concerning game state consistency among the participating peers, vulnerability to cheating, and increased bandwidth usage by each peer. There are some proposals that try to minimize these problems. One of them is presented in \cite{schiele2007rpp}, where it is suggested the division of the virtual simulated game environment into regions, and for each region, a peer is chosen to be its coordinator. The function of this peer will be manage the players' interest. It will check who really needs each state update, so each player does not receive every update packet. This way, the peers' bandwidth usage is reduced. However, it will be still higher than with a server performing all the broadcasts. In the peer-to-peer model, each player usually must send updates to more than one other player. Besides, the chosen peer must be trustable to be the region coordinator.

@@

In \cite{zou2001egt}, multicast-based interest management schemes are also considered. A comparison is made between cell-based grouping, where each group is associated with a cell of a grid that represents the virtual environment, and grouping based on objects, where there is a multicast group for each object. It was found that there is a tradeoff between the cost of multicast group formation control messages and the cost of the state updates themselves.

@@

In \cite{morse1996iml}, it is made a survey of systems that use the interest management technique, describing the purpose, scope and what criteria are used by the management scheme employed by each one of them. It is then presented a taxonomy of such systems, classifying them according to: communication model, filtering focus and domain of responsibility. The model may be \textit{unicast}, \textit{multicast} or \textit{broadcast}. The filtering focus refers to what characteristics are observed for each object to perform the filtering: they may be \textit{intrinsic}, such as the values of the object's attributes (e.g. the exact coordinates of its location), or \textit{extrinsic}, such as which multicast group it is subscribed to. Finally, the domain of responsibility assigned to an interest manager, which checks for whom each state update is relevant, can be \textit{static} or \textit{dynamic}. For instance, if an interest manager is assigned to a fixed area of the virtual environment, its domain of responsibility is static, but if he controls an area that may increase or decrease in size, its domain of responsibility is dynamic.

@@

Taking into consideration that the multicast communication model is not widely supported in the Internet \cite{elsayed2004alm}, %for technical and marketing reasons as well,
 in this work we opted to follow the unicast model. Each broadcast consists then of a set of unicast transmissions, one for each destination. Moreover, intrinsic filtering and dynamic domain of responsibility were used, in order to increase the precision of the interest management and consequently achieve a greater reduction in the state update traffic \cite{morse1996iml}.

@@

This work is part of the P2PSE \cite{crippa2007pps} project, which is currently based on a peer-to-peer distribution, but still looking for an ideal model. Hence, we propose here an alternative that consists of a virtual environment partitioned into regions - each managed by a different server - on which the \malg\ algorithm is based. The regions are contiguous, exploring the locality of the players' avatars. Thus, avatars next to each other will probably be located in the same region and, therefore, their clients will tend to be connected to the same server, so that their interaction is faster (Figure \ref{partitions}). Situations in which two players interacting with each other are connected to different servers imply a higher bandwidth usage, as it is required some sort of negotiation between the servers to which the different players are connected, in order to maintain the states of the simulation identical in both servers. Also, the hop count of each message between the clients of these players is increased, since there is one more intermediary.

@@

% use section* for acknowledgement
\section*{Acknowledgment}

The development of this work was supported by the Funding for Studies and Projects (FINEP), a brazilian research agency, through the P2PSE project (P2PSE-5849-1).
